Calendar-based graphics for visualizing people's daily schedules

### 2017-09-07 main.Rmd
mv sugrrants citation to the first paragraph of intro

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualizing people's daily schedules"
  plain: "Calendar-based graphics for visualizing people's daily schedules"
abstract: >
  Calendars are broadly used in society to display temporal information, and events. This paper describes a new \proglang{R} package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout. The function \code{frame\_calendar} uses linear algebra on the date variable to re-structure data into a format lending itself to calendar layouts. The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the \proglang{R} package \pkg{sugrrants}.
keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \pkg{ggplot2} [@R-ggplot2]. The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite. The methods are implemented in a new package called \pkg{sugrrants}.

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series. The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package [@R-sugrrants] using \proglang{R} [@R-base].

@@

>>>>

---
author:
  - name: Earo Wang
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{earo.wang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{dicook@monash.edu}
  - name: Rob J Hyndman
    affiliation: Monash University
    address: >
      Department of Econometrics and Business Statistics,
      Monash University, VIC 3800
      Australia
    email: \email{rob.hyndman@monash.edu}
title:
  formatted: "Calendar-based graphics for visualizing people's daily schedules"
  plain: "Calendar-based graphics for visualizing people's daily schedules"
abstract: >
  Calendars are broadly used in society to display temporal information, and events. This paper describes a new \proglang{R} package with functionality to organize and display temporal data, collected on sub-daily resolution, into a calendar layout. The function \code{frame\_calendar} uses linear algebra on the date variable to restructure data into a format lending itself to calendar layouts. The user can apply the grammar of graphics to create plots inside each calendar cell, and thus the displays synchronize neatly with \pkg{ggplot2} graphics. The motivating application is studying pedestrian behavior in Melbourne, Australia, based on counts which are captured at hourly intervals by sensors scattered around the city. Faceting by the usual features such as day and month, was insufficient to examine the behavior. Making displays on a monthly calendar format helps to understand pedestrian patterns relative to events such as work days, weekends, holidays, and special events. The layout algorithm has several format options and variations. It is implemented in the \proglang{R} package \pkg{sugrrants}.
keywords:
  # at least one keyword must be supplied
  formatted: [data visualization, statistical graphics, time series, grammar of graphics, "\\proglang{R}"]
  plain:     [data visualization, statistical graphics, time series, grammar of graphics, R]
preamble: >
  \usepackage{amsmath}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
  citation_package: biblatex
bibliography: ["references.bib", "packages.bib"]
---

@@

We develop a method for organizing and visualizing temporal data, collected at sub-daily intervals, into a calendar layout. The calendar format is created using linear algebra, giving a restructuring of the data, that can then be piped into grammar of graphics definitions of plots, as used in \pkg{ggplot2} [@R-ggplot2]. The data restructuring approach is consistent with the tidy data principles available in the \pkg{tidyverse} [@R-tidyverse] suite. The methods are implemented in a new package called \pkg{sugrrants} [@R-sugrrants] .

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series. The proposed algorithm has been implemented in the \code{frame_calendar} function in the \pkg{sugrrants} package using \proglang{R} [@R-base].

@@

\newpage

### 2017-09-07 main.tex
mv sugrrants citation to the first paragraph of intro

\Abstract{
Calendars are broadly used in society to display temporal information,
and events. This paper describes a new \proglang{R} package with
functionality to organize and display temporal data, collected on
sub-daily resolution, into a calendar layout. The function
\code{frame\_calendar} uses linear algebra on the date variable to
re-structure data into a format lending itself to calendar layouts. The
user can apply the grammar of graphics to create plots inside each
calendar cell, and thus the displays synchronize neatly with
\pkg{ggplot2} graphics. The motivating application is studying
pedestrian behavior in Melbourne, Australia, based on counts which are
captured at hourly intervals by sensors scattered around the city.
Faceting by the usual features such as day and month, was insufficient
to examine the behavior. Making displays on a monthly calendar format
helps to understand pedestrian patterns relative to events such as work
days, weekends, holidays, and special events. The layout algorithm has
several format options and variations. It is implemented in the
\proglang{R} package \pkg{sugrrants}.
}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, giving a restructuring of the
data, that can then be piped into grammar of graphics definitions of
plots, as used in \pkg{ggplot2} \citep{R-ggplot2}. The data
restructuring approach is consistent with the tidy data principles
available in the \pkg{tidyverse} \citep{R-tidyverse} suite. The methods
are implemented in a new package called \pkg{sugrrants}.

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The
approach is developed with three conditions in mind: (1) to display
time-of-day variation in addition to longer temporal components such as
day-of-week and day-of-year; (2) to incorporate line graphs and other
types of glyphs into the graphical toolkit for the calendar layout; (3)
to enable overlaying plots consisting of multiple time series. The
proposed algorithm has been implemented in the \code{frame_calendar}
function in the \pkg{sugrrants} package \citep{R-sugrrants} using
\proglang{R} \citep{R-base}.

>>>>

\Abstract{
Calendars are broadly used in society to display temporal information,
and events. This paper describes a new \proglang{R} package with
functionality to organize and display temporal data, collected on
sub-daily resolution, into a calendar layout. The function
\code{frame\_calendar} uses linear algebra on the date variable to
restructure data into a format lending itself to calendar layouts. The
user can apply the grammar of graphics to create plots inside each
calendar cell, and thus the displays synchronize neatly with
\pkg{ggplot2} graphics. The motivating application is studying
pedestrian behavior in Melbourne, Australia, based on counts which are
captured at hourly intervals by sensors scattered around the city.
Faceting by the usual features such as day and month, was insufficient
to examine the behavior. Making displays on a monthly calendar format
helps to understand pedestrian patterns relative to events such as work
days, weekends, holidays, and special events. The layout algorithm has
several format options and variations. It is implemented in the
\proglang{R} package \pkg{sugrrants}.
}

@@

We develop a method for organizing and visualizing temporal data,
collected at sub-daily intervals, into a calendar layout. The calendar
format is created using linear algebra, giving a restructuring of the
data, that can then be piped into grammar of graphics definitions of
plots, as used in \pkg{ggplot2} \citep{R-ggplot2}. The data
restructuring approach is consistent with the tidy data principles
available in the \pkg{tidyverse} \citep{R-tidyverse} suite. The methods
are implemented in a new package called \pkg{sugrrants}
\citep{R-sugrrants} .

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The
approach is developed with three conditions in mind: (1) to display
time-of-day variation in addition to longer temporal components such as
day-of-week and day-of-year; (2) to incorporate line graphs and other
types of glyphs into the graphical toolkit for the calendar layout; (3)
to enable overlaying plots consisting of multiple time series. The
proposed algorithm has been implemented in the \code{frame_calendar}
function in the \pkg{sugrrants} package using \proglang{R}
\citep{R-base}.
